On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-06-02 manuscript.Rmd
Nico revisions added

\
The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest because of its simplicity is the method where leaves are detached from the plant, their weight loss monitored over time, and expressed as the minimum conductance (g~min~). This method (mass loss of detached leaves, MLD) aims to simulate field conditions during severe drought, when water supply to the leaf has practically ceased, although it proceeds much more quickly (typically 0.5-2 days, compared to weeks or even months in field conditions). This measurement also includes both surfaces of the leaf - not just the adaxial side. A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.

@@

The minimum leaf conductance includes two pathways: across the cuticle, and through the (potentially incompletely closed) stomata. Only a few studies have directly quantified the stomatal component of minimum conductance, in contrast to the wealth of information on the cuticular component. A detailed study of *Hedera helix*  concluded that 35% of water loss occurred across the stomatal pores (though closed), and 65% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004]. For ten *Sorghum* genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~. For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky, and likely incompletely closed. However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more permeable than the non-stomatal side, thus providing another potential explanation for this difference. Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies among species.

@@

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data [though other mechanisms can also be employed to simulate this pattern, see @dewar2018]. When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

The previous section discussed and compared methods for more or less direct measurements of various definitions of minimum conductance. The approach taken by the majority of vegetation models is however very different. Usually, g~0~ (for use in Eq. 1) is estimated from regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor (a combination of measured photosynthesis rate, air humidity, and CO~2~ concentration). The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species, presumably because unreliable estimates were obtained), b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values. @barnard2013 also mentioned the difficulty of fitting Eq. 1 to estimate g~0~, but did not present specific details. The following analysis builds on their work by demonstrating statistical uncertainties.

@@

\
This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see e.g. @misson2004]. Another advantage is that independent estimates of g~0~ can be used, not those obtained via regression, which produces the undesirable correlation with estimates of g~1~. The above formulation is yet to be tested against data, but we propose this test to be done with data from drought and non-drought conditions.

>>>>

\
A comprehensive review of g~min~ with the goal to improve functional model representation is made difficult by the fact that literature data arise from various methods, and represent distinct processes. The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest because of its simplicity is the method where leaves are detached from the plant, their weight loss monitored over time, and expressed as the minimum conductance (g~min~). This method (mass loss of detached leaves, MLD) aims to simulate field conditions during severe drought, when water supply to the leaf has practically ceased, although it proceeds much more quickly (typically 0.5-2 days, compared to weeks or even months in field conditions). This measurement also includes both surfaces of the leaf - not just the adaxial side. A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.

@@

The minimum leaf conductance includes two pathways: across the cuticle, and through the (potentially incompletely closed) stomata. Only a few studies have directly quantified the stomatal component of minimum conductance, in contrast to the wealth of information on the cuticular component. A detailed study of *Hedera helix*  concluded that 35% of water loss occurred across the stomatal pores (though closed), and 65% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004]. For ten *Sorghum* genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~. For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky, and likely incompletely closed. However, @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more permeable than the non-stomatal side, thus providing another potential explanation for this difference. Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies among species.

@@

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data [though other mechanisms can also be employed to simulate this pattern, see @dewar2018]. When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~i~. To see this point, we can rearrange Eq. 1 to give,

@@

The previous section discussed and compared methods for more or less direct measurements of minimum conductance according to various definitions. The approach taken by the majority of vegetation models is however very different. Usually, g~0~ (for use in Eq. 1) is estimated from regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor (a combination of measured photosynthesis rate, air humidity, and CO~2~ concentration). The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species, presumably because unreliable estimates were obtained), b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values. @barnard2013 also mentioned the difficulty of fitting Eq. 1 to estimate g~0~, but did not present specific details. The following analysis builds on their work by demonstrating statistical uncertainties.

@@

\
This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see e.g. @misson2004] - as long as the right-hand term (including g~1~) is reduced under drought as well [@zhou2013]. Another advantage is that independent estimates of g~0~ can be used, not those obtained via regression, which produces the undesirable correlation with estimates of g~1~. The above formulation is yet to be tested against data, but we propose this test to be done with data from drought and non-drought conditions.
